How [preposition inanimate] becomes [\tsc{r}-pronoun postposition]\\ On \tit{waarmee} vs. \tit{met wat} in Dutch

### 2020-01-14 paper.tex
during colloquium
added sigma in trees

\title{No preposition with inanimate but an \tsc{r}-pronoun with postposition\\
On \tit{waarmee} vs. \tit{met wat} in Dutch}
\author{Fenna Bergsma}
\date{\today}

@@

English has a sentence. This is ungrammatical in Dutch. We need an \tsc{r}-pronoun. The term \tsc{r}-pronoun \citep{riemsdijk1978} refers to a set of nominal elements that can strand prepositions in Dutch (and German). This is noteworthy because Dutch is a normally non-preposition stranding language. \tsc{R}-pronouns morphologically consist of a combination of a preposition and locative pronoun.
In the example \ref{ex:klimerop} \tit{op} `on' is the preposition and \tit{'r} `there' is the locative pronoun.\footnote{The \tsc{r}-pronoun in \ref{ex:klimerop} can be written as \tit{er}, \tit{der} and \tit{'r} and pronounced as respectively /ɛr/, /dər/ or /ər/ cf. \citet{wesseling2018}. For this paper I consider them to be allomorphs. In my examples I use \tit{'r}.}

@@

This is a free relative construction in which the two predicates (the one in the main clause and the one in the embedded clause) combine with two different cases (i.e. the case requirements do not match). I illustrate this in \ref{ex:gekochtwaarmee}. The predicate in the embedded clause, \tit{schildert} `paint', combines with an instrumental PP. The predicate in the main clause clause, \tit{gekocht} `bought' combines with an accusative DP. The \tsc{r}-pronoun \tit{waarmee} `with what' is used here.\footnote{In this example, \tit{waar} `where' takes \tit{mee} `with' to the left edge of the embedded clause. It is also possible for \tit{mee} `with' to be stranded, and \tit{waar} `where' to be moved to the left edge of the embedded clause on its own.

@@

\tit{Waarmee} `with what' and not \tit{met wat} `with what' does not only appear in wh-questions, but also in other contexts. \ref{ex:headed} gives an example of a headed relative, and \ref{ex:headless} shows a free relative in which both predicates combine with an instrumental PP. The use of \tit{met wat} `with what' is ungrammatical in both contexts, and \tit{waarmee} `with what' is used.

@@

In this section I showed that ʀ-pronouns are expected in combinations with prepositions. This means that the use of \tit{waarmee} `with what' \ref{ex:grammaticalwaar} is not surprising. Something that is surprising is the use of \tit{met wat} `with what' in \ref{ex:grammaticalwat}, and this is the example something more needs to be said about. In the remainder of this section I argue that this `something more' is that the instrumental PP in \ref{ex:grammaticalwat} does not form a proper constituent, i.e. it is not a constituent to the exclusion of any other elements. The other side of the coin is that constructions with \tsc{r}-pronouns contain a PP that does form a proper constituent.

Below I repeat the examples with instrumental PPs I discuss so far in this paper.

@@

In each of these examples the instrumental PP forms a constituent at a certain point in the derivation. In \ref{ex:const1}, the PP forms a proper constituent in the surface order, as shown in \ref{ex:const1stage}. In \ref{ex:const2}, the PP forms a proper constituent before wh- and V2- movement, shown in \ref{ex:const2stage}.
The structure in \ref{ex:const3stage} represents a stage in the derivation of the embedded clauses in \ref{ex:const3} and \ref{ex:const4}.\footnote{I assume that the antecedent in a free relative is a phonologically empty element, in line with cf. \cite{bresnan1978a,groos1981,himmelreich2017}. Under that view, \ref{ex:const3} and \ref{ex:const4} are identical, except for that in \ref{ex:const4} modifies phonologically empty element.}
Again, in the stage, which comes before relative movement of the pronoun to the left periphery of the relative clause, the PP forms a proper constituent.

@@

Let me now show how this applies to the examples with the mismatching free relatives. The two predicates I used in the free relatives are \tit{kopen} `to buy' and \tit{schilderen}  `to paint'. \tit{Kopen} `to buy' takes an accusative DP as its object, illustrated in \ref{ex:kopen}. \tit{Schilderen} `to paint' can take an instrumental PP as its object., shown in \ref{ex:schilderen}.\footnote{\tit{Schilderen} also optionally takes an (accusative) object, but I am focussing on the instrumental object here.}

@@

I repeat the mismatching free relative in which \tit{waarmee} `met wat' appears in \ref{ex:mismatchwaarmee}. The predicate \tit{schildert} `paints' combines in the embedded clause with the instrumental object. The PP forms a proper constituent within the embedded clause, and the PP can be realized as the \tsc{r}-pronoun \tit{waarmee} `with what'.\footnote{I assume that the embedded clause modifies a phonologically empty noun here which is the DP argument of \tit{gekocht} `bought'.}

@@

Next, we arrive at the mismatching free relative in which \tit{waarmee} `with what' cannot be used, but \tit{met wat} `with what' appears. The embedded clause predicate \tit{gekocht} `bought' combines with an accusative DP. The accusative object of a verb is always \tit{wat} `what', as I showed in \ref{ex:wat}. The instrumental PP only comes into the picture in the main clause, when \tit{schilder} `paint' combines with an instrumental PP. At no point in the derivation does the PP form a proper constituent, and \tit{waarmee} `with what' does not surface.

@@

\subsection{Differences: \tit{'r} vs. \tit{'t} and \tit{met} vs. \tit{-mee}}

@@

 \ex.
 \a. \begin{forest} boom
 [\tsc{acc}P
     [\tsc{f}2]
     [\tsc{nom}P
         [\tsc{f}1]
         [\tsc{thingP}
             [\tsc{thing}, roof]
         ]
     ]
 ]
 \end{forest}\label{ex:tacc}
 \b. \begin{forest} boom
 [\tsc{nom}P
     [\tsc{f}1]
     [\tsc{thingP}
         [\tsc{thing}, roof]
     ]
 ]
 \end{forest}\label{ex:tnom}

@@

The syntactic structure in \ref{ex:entryr} only has \tsc{loc} as an unused feature, whereas \ref{ex:entryt} has Σ up to \tsc{f3}.

@@

\ex. \begin{forest} boom
[\tsc{com}P
    [\tsc{f}5]
    [\tsc{insP}
        [\tsc{f}4]
        [\tsc{dat}P
            [\tsc{f}3]
            [\tsc{acc}P
                [\tsc{f}2]
                [\tsc{nom}P
                    [\tsc{f}1]
                ]
            ]
        ]
    ]
]
{\draw (.east) node[right]{⇔ \tit{mee}}; }
\end{forest}\label{ex:entrymee}

@@

\ex. [ [ [ [ [ [ \tsc{thing} ] \tsc{deix} ] W ] \tsc{f1} ] \tsc{f2} ] \tsc{f3} \tsc{f4} ]\label{ex:fseq}
\z.

@@

\ex. \begin{forest} boom
[\tsc{nom}P
   [\tsc{f}1]
   [\tsc{thingP}
       [\tsc{thing}, roof]
   ]
]
{\draw (.east) node[right]{⇒ \tit{'t}}; }
\end{forest}\label{ex:thingf1}

@@

\ex. \begin{forest} boom
[\tsc{accP}
   [\tsc{f}2]
   [\tsc{nom}P
       [\tsc{f}1]
       [\tsc{thingP}
           [\tsc{thing}, roof]
       ]
   ]
]
{\draw (.east) node[right]{⇒ \tit{'t}}; }
\end{forest}

@@

\ex. \begin{forest} boom
[\tsc{datP}
    [\tsc{f}3]
    [\tsc{accP}
       [\tsc{f}2]
       [\tsc{nom}P
           [\tsc{f}1]
           [\tsc{thingP}
               [\tsc{thing}, roof]
           ]
       ]
    ]
]
{\draw (.east) node[right]{⇒ \tit{'t}}; }
\end{forest}

@@

\ex.
\a. \begin{forest} boom
[\tsc{insP}
    [\tsc{f}4]
    [\tsc{datP}
        [\tsc{f}3]
        [\tsc{accP}
           [\tsc{f}2]
           [\tsc{nom}P
               [\tsc{f}1]
               [\tsc{thingP}
                   [\tsc{thing}, roof]
               ]
           ]
        ]
    ]
]
{\draw (.east) node[right]{⇒ }; }
\end{forest}\label{ex:f4no}
\b. \begin{forest} boom
[\tsc{datP}
    [\tsc{datP}
        [\tsc{f}3]
        [\tsc{accP}
           [\tsc{f}2]
           [\tsc{nom}P
               [\tsc{f}1]
               [\tsc{thingP}
                   [\tsc{thing}, roof]
               ]
           ]
        ]
    ]
    {\draw (.east) node[right]{⇒ \tit{'t}}; }
    [\tsc{insP}
        [\tsc{f}4]
    ]
    {\draw (.east) node[right]{⇒ }; }
]
\end{forest}\label{ex:f4comp}

@@

\begin{forest} boom
[\tsc{accP}
    [\tsc{accP}
       [\tsc{f}2]
       [\tsc{nom}P
           [\tsc{f}1]
           [\tsc{thingP}
               [\tsc{thing}, roof]
           ]
       ]
    ]
    {\draw (.east) node[right]{⇒ \tit{'t}}; }
    [\tsc{datP}
        [\tsc{f}3]
    ]
    {\draw (.east) node[right]{⇒ }; }
]
\end{forest}\label{ex:f3comp}

Backtracking proceeds further, into the cycle in which \tsc{f2} was merged. Again there is no specifier to be moved, and complement movement does not give any results.

The derivation ends up backtracking into the cycle in which \tsc{f1} was merged. At this stage, \tsc{thing} was realized as \tit{'r}. Here there is again no specifier, but complement movement provides a structure that is fit for a lexical entry, namely: \ref{ex:entrymee}.

\ex. \begin{forest} boom
[\tsc{nomP}
   [\tsc{thingP}
       [\tsc{thing}, roof]
   ]
   {\draw (.east) node[right]{⇒ \tit{'r}}; }
   [\tsc{nom}P
       [\tsc{f}1]
   ]
   {\draw (.east) node[right]{⇒ \tit{mee}}; }
]
\end{forest}

The feature \tsc{f}2 is merged again. The structure cannot be spelled out as a whole, so spec-to-spec movement is tried. In that configuration \tsc{f2} can be realized together with \tsc{f1} as \tit{mee}.

\ex. \begin{forest} boom
[\tsc{accP}
   [\tsc{thingP}
       [\tsc{thing}, roof]
   ]
   {\draw (.east) node[right]{⇒ \tit{'r}}; }
   [\tsc{acc}P
       [\tsc{f}2]
       [\tsc{nom}P
           [\tsc{f}1]
       ]
   ]
   {\draw (.east) node[right]{⇒ \tit{mee}}; }
 ]
\end{forest}

The same happens for \tsc{f3} and \tsc{f4}, and the complete structure is realizes as shown in \ref{ex:spellout'rmee}.

\ex. \begin{forest} boom
[\tsc{insP}
    [\tsc{thingP}
       [\tsc{thing}, roof]
    ]
    {\draw (.east) node[right]{⇒ \tit{'r}}; }
    [\tsc{insP}
       [\tsc{f}4]
       [\tsc{dat}P
           [\tsc{f}3]
           [\tsc{acc}P
               [\tsc{f}2]
               [\tsc{nom}P
                   [\tsc{f}1]
               ]
           ]
       ]
    ]
    {\draw (.east) node[right]{⇒ \tit{mee}}; }
]
\end{forest}\label{ex:spellout'rmee}

@@

\ex. \begin{forest} boom
[\tsc{WP}
    [\tsc{WP}
        [\tsc{W}, roof]
    ]
    {\draw (.east) node[right]{⇒ \tit{w}}; }
    [\tsc{deixP}
        [\tsc{deixP}
            [\tsc{deix}, roof]
        ]
        {\draw (.east) node[right]{⇒ \tit{a}}; }
        [\tit{insP}
            [\tsc{thingP}
               [\tsc{thing}, roof]
            ]
            {\draw (.east) node[right]{⇒ \tit{'r}}; }
            [\tsc{insP}
               [\tsc{f}4]
               [\tsc{dat}P
                   [\tsc{f}3]
                   [\tsc{acc}P
                       [\tsc{f}2]
                       [\tsc{nom}P
                           [\tsc{f}1]
                       ]
                   ]
               ]
            ]
            {\draw (.east) node[right]{⇒ \tit{mee}}; }
        ]
    ]
]
\end{forest}\label{ex:spelloutwaarmee}

@@

\ex. \begin{forest} boom
[CP
    [CP
        [...,roof]
    ]
    [\tsc{WP}
        [\tsc{WP}
            [\tsc{W}, roof]
        ]
        {\draw (.east) node[right]{⇒ \tit{w}}; }
        [\tsc{deixP}
            [\tsc{deixP}
                [\tsc{deix}, roof]
            ]
            {\draw (.east) node[right]{⇒ \tit{a}}; }
            [\tsc{accP}
               [\tsc{f}2]
               [\tsc{nom}P
                   [\tsc{f}1]
                   [\tsc{thingP}
                       [\tsc{thing}, roof]
                   ]
               ]
            ]
            {\draw (.east) node[right]{⇒ \tit{'t}}; }
        ]
    ]
]
\end{forest}

If a dative feature is merged here, it will not be able to be spelled out. The only option to realize the features is to form to spawn a new derivation, to create a complex specifier. And that is exactly what forming the preposition \tit{met} `with' is.

>>>>

\title{How [preposition inanimate] becomes [\tsc{r}-pronoun postposition]\\
On \tit{waarmee} vs. \tit{met wat} in Dutch}
\author{Fenna Bergsma}
\date{\today}

@@

English has a sentence. This is ungrammatical in Dutch. We need an \tsc{r}-pronoun. The term \tsc{r}-pronoun \citep{riemsdijk1978} refers to a set of nominal elements that can strand prepositions in Dutch (and German). This is noteworthy because Dutch is a normally non-preposition stranding language.
In the example \ref{ex:klimerop} \tit{op} `on' is the preposition and \tit{'r} `there' is the locative pronoun.\footnote{The \tsc{r}-pronoun \tit{'r} in \ref{ex:klimerop} can be written as \tit{er}, \tit{der} and \tit{'r} and pronounced as respectively /ɛr/, /dər/ or /ər/. As far as I am aware, there is no clear meaning difference between these forms. See \citet{wesseling2018} for discussion. In my examples I use \tit{'r}, but the other two forms fit just as well.}

@@

This is a free relative construction in which the two predicates (the one in the main clause and the one in the embedded clause) combine with two different cases (i.e. the case requirements do not match). I illustrate this in \ref{ex:gekochtwaarmee}. The predicate in the embedded clause, \tit{schildert} `paint', combines with an instrumental object. The predicate in the main clause clause, \tit{gekocht} `bought' combines with an accusative DP. The \tsc{r}-pronoun \tit{waarmee} `with what' is used here.\footnote{In this example, \tit{waar} `where' takes \tit{mee} `with' to the left edge of the embedded clause. It is also possible for \tit{mee} `with' to be stranded, and \tit{waar} `where' to be moved to the left edge of the embedded clause on its own.

@@

\tit{Waarmee} `with what' and not \tit{met wat} `with what' does not only appear in wh-questions, but also in other contexts. \ref{ex:headed} gives an example of a headed relative, and \ref{ex:headless} shows a free relative in which both predicates combine with an instrumental object. The use of \tit{met wat} `with what' is ungrammatical in both contexts, and \tit{waarmee} `with what' is used.

@@

In this section I showed that ʀ-pronouns are expected in combinations with prepositions. This means that the use of \tit{waarmee} `with what' \ref{ex:grammaticalwaar} is not surprising. Something that is surprising is the use of \tit{met wat} `with what' in \ref{ex:grammaticalwat}, and this is the example something more needs to be said about. In the remainder of this section I argue that this `something more' is that the instrumental object in \ref{ex:grammaticalwat} does not form a proper constituent, i.e. it is not a constituent to the exclusion of any other elements. The other side of the coin is that constructions with \tsc{r}-pronouns contain a object that does form a proper constituent.

Below I repeat the examples with instrumentals I discussed so far in this paper.

@@

In each of these examples the instrumental object forms a constituent at a certain point in the derivation. In \ref{ex:const1}, the instrumental object forms a proper constituent in the surface order, as shown in \ref{ex:const1stage}. In \ref{ex:const2}, the instrumental object forms a proper constituent before wh- and V2- movement, shown in \ref{ex:const2stage}.
The structure in \ref{ex:const3stage} represents a stage in the derivation of the embedded clauses in \ref{ex:const3} and \ref{ex:const4}.\footnote{I assume that the antecedent in a free relative is a phonologically empty element, in line with cf. \cite{bresnan1978a,groos1981,himmelreich2017}. Under that view, \ref{ex:const3} and \ref{ex:const4} are identical, except for that in \ref{ex:const4} modifies phonologically empty element.}
Again, in the stage, which comes before relative movement of the pronoun to the left periphery of the relative clause, the instrumental object forms a proper constituent.

@@

Let me now show how this applies to the examples with the mismatching free relatives. The two predicates I used in the free relatives are \tit{kopen} `to buy' and \tit{schilderen}  `to paint'. \tit{Kopen} `to buy' takes an accusative DP as its object, illustrated in \ref{ex:kopen}. \tit{Schilderen} `to paint' can take an instrumental as its object., shown in \ref{ex:schilderen}.\footnote{\tit{Schilderen} also optionally takes an (accusative) object, but I am focussing on the instrumental object here.}

@@

I repeat the mismatching free relative in which \tit{waarmee} `met wat' appears in \ref{ex:mismatchwaarmee}. The predicate \tit{schildert} `paints' combines in the embedded clause with the instrumental object. The instrumental object forms a proper constituent within the embedded clause, and the it can be realized as the \tsc{r}-pronoun and postposition \tit{waarmee} `with what'.\footnote{I assume that the embedded clause modifies a phonologically empty noun here which is the DP argument of \tit{gekocht} `bought'.}

@@

Next, we arrive at the mismatching free relative in which \tit{waarmee} `with what' cannot be used, but \tit{met wat} `with what' appears. The embedded clause predicate \tit{gekocht} `bought' combines with an accusative DP. The accusative object of a verb is always \tit{wat} `what', as I showed in \ref{ex:wat}. The instrumental only comes into the picture in the main clause, when \tit{schilder} `paint' combines with an instrumental object. At no point in the derivation does the instrumental object form a proper constituent, and \tit{waarmee} `with what' does not surface.

@@

\subsection{Differences: \tit{'rmee} vs. \tit{met 't}}

@@

 \ex.
 \a. \begin{forest} boom
 [\tsc{acc}P
     [\tsc{f}2]
     [\tsc{nom}P
         [\tsc{f}1]
         [ΣP
             [Σ]
             [\tsc{thingP}
                 [\tsc{thing}, roof]
             ]
         ]
     ]
 ]
 \end{forest}\label{ex:tacc}
 \b. \begin{forest} boom
 [\tsc{nom}P
     [\tsc{f}1]
     [ΣP
         [Σ]
         [\tsc{thingP}
             [\tsc{thing}, roof]
         ]
     ]
 ]
 \end{forest}\label{ex:tnom}

@@

The syntactic structure in \ref{ex:entryr} only has \tsc{loc} as an unused feature, whereas in \ref{ex:entryt} Σ up to \tsc{f3} remain unused.

@@

\ex. \begin{forest} boom
[\tsc{com}P
    [\tsc{f}5]
    [\tsc{insP}
        [\tsc{f}4]
        [\tsc{dat}P
            [\tsc{f}3]
            [\tsc{acc}P
                [\tsc{f}2]
                [\tsc{nom}P
                    [\tsc{f}1]
                    [ΣP
                        [Σ]
                    ]
                ]
            ]
        ]
    ]
]
{\draw (.east) node[right]{⇔ \tit{mee}}; }
\end{forest}\label{ex:entrymee}

@@

\ex. [ [ [ [ [ [ [ \tsc{thing} ] \tsc{deix} ] W ] \tsc{f1} ] \tsc{f2} ] \tsc{f3} ] \tsc{f4} ]\label{ex:fseq}
\z.

@@

\ex. \begin{forest} boom
[\tsc{nom}P
   [\tsc{f}1]
   [ΣP
       [Σ]
       [\tsc{thingP}
           [\tsc{thing}, roof]
       ]
   ]
]
{\draw (.east) node[right]{⇒ \tit{'t}}; }
\end{forest}\label{ex:thingf1}

@@

\ex. \begin{forest} boom
[\tsc{accP}
   [\tsc{f}2]
   [\tsc{nom}P
       [\tsc{f}1]
       [ΣP
           [Σ]
           [\tsc{thingP}
               [\tsc{thing}, roof]
           ]
       ]
   ]
]
{\draw (.east) node[right]{⇒ \tit{'t}}; }
\end{forest}

@@

\ex. \begin{forest} boom
[\tsc{datP}
    [\tsc{f}3]
    [\tsc{accP}
       [\tsc{f}2]
       [\tsc{nom}P
           [\tsc{f}1]
           [ΣP
               [Σ]
               [\tsc{thingP}
                   [\tsc{thing}, roof]
               ]
           ]
       ]
    ]
]
{\draw (.east) node[right]{⇒ \tit{'t}}; }
\end{forest}

@@

\ex.
\a. \begin{forest} boom
[\tsc{insP}
    [\tsc{f}4]
    [\tsc{datP}
        [\tsc{f}3]
        [\tsc{accP}
           [\tsc{f}2]
           [\tsc{nom}P
               [\tsc{f}1]
               [ΣP
                   [Σ]
                   [\tsc{thingP}
                       [\tsc{thing}, roof]
                   ]
               ]
           ]
        ]
    ]
]
{\draw (.east) node[right]{⇒ }; }
\end{forest}\label{ex:f4no}
\b. \begin{forest} boom
[\tsc{datP}
    [\tsc{datP}
        [\tsc{f}3]
        [\tsc{accP}
           [\tsc{f}2]
           [\tsc{nom}P
               [\tsc{f}1]
               [ΣP
                   [Σ]
                   [\tsc{thingP}
                       [\tsc{thing}, roof]
                   ]
               ]
           ]
        ]
    ]
    {\draw (.east) node[right]{⇒ \tit{'t}}; }
    [\tsc{insP}
        [\tsc{f}4]
    ]
    {\draw (.east) node[right]{⇒ }; }
]
\end{forest}\label{ex:f4comp}

@@

\begin{forest} boom
[\tsc{accP}
    [\tsc{accP}
       [\tsc{f}2]
       [\tsc{nom}P
           [\tsc{f}1]
           [ΣP
               [Σ]
               [\tsc{thingP}
                   [\tsc{thing}, roof]
               ]
           ]
       ]
    ]
    {\draw (.east) node[right]{⇒ \tit{'t}}; }
    [\tsc{datP}
        [\tsc{f}3]
    ]
    {\draw (.east) node[right]{⇒ }; }
]
\end{forest}\label{ex:f3comp}

Backtracking proceeds further, into the cycle in which \tsc{f1} was merged. Again there is no specifier to be moved, and complement movement does not give any results.

The derivation ends up backtracking into the cycle in which Σ was merged. At this stage, \tsc{thing} was realized as \tit{'r}. Here there is again no specifier, but complement movement provides a structure that is fit for a lexical entry, namely: \ref{ex:entrymee}.

\ex. \begin{forest} boom
[ΣP
   [\tsc{thingP}
       [\tsc{thing}, roof]
   ]
   {\draw (.east) node[right]{⇒ \tit{'r}}; }
   [ΣP
       [Σ]
   ]
   {\draw (.east) node[right]{⇒ \tit{mee}}; }
]
\end{forest}

The feature \tsc{f}1 is merged again. The structure cannot be spelled out as a whole, so spec-to-spec movement is tried. In that configuration \tsc{f1} can be realized together with Σ as \tit{mee}.

\ex. \begin{forest} boom
[\tsc{nomP}
   [\tsc{thingP}
       [\tsc{thing}, roof]
   ]
   {\draw (.east) node[right]{⇒ \tit{'r}}; }
   [\tsc{nom}P
       [\tsc{f}1]
       [ΣP
           [Σ]
       ]
   ]
   {\draw (.east) node[right]{⇒ \tit{mee}}; }
 ]
\end{forest}

The same happens for \tsc{f2}, \tsc{f3} and \tsc{f4}, and the complete structure is realizes as shown in \ref{ex:spellout'rmee}.

\ex. \begin{forest} boom
[\tsc{insP}
    [\tsc{thingP}
       [\tsc{thing}, roof]
    ]
    {\draw (.east) node[right]{⇒ \tit{'r}}; }
    [\tsc{insP}
       [\tsc{f}4]
       [\tsc{dat}P
           [\tsc{f}3]
           [\tsc{acc}P
               [\tsc{f}2]
               [\tsc{nom}P
                   [\tsc{f}1]
                   [ΣP
                       [Σ]
                   ]
               ]
           ]
       ]
    ]
    {\draw (.east) node[right]{⇒ \tit{mee}}; }
]
\end{forest}\label{ex:spellout'rmee}

@@

\ex. \begin{forest} boom
[\tsc{WP}
    [\tsc{WP}
        [\tsc{W}, roof]
    ]
    {\draw (.east) node[right]{⇒ \tit{w}}; }
    [\tsc{deixP}
        [\tsc{deixP}
            [\tsc{deix}, roof]
        ]
        {\draw (.east) node[right]{⇒ \tit{a}}; }
        [\tit{insP}
            [\tsc{thingP}
               [\tsc{thing}, roof]
            ]
            {\draw (.east) node[right]{⇒ \tit{'r}}; }
            [\tsc{insP}
               [\tsc{f}4]
               [\tsc{dat}P
                   [\tsc{f}3]
                   [\tsc{acc}P
                       [\tsc{f}2]
                       [\tsc{nom}P
                           [\tsc{f}1]
                           [ΣP
                               [Σ]
                           ]
                       ]
                   ]
               ]
            ]
            {\draw (.east) node[right]{⇒ \tit{mee}}; }
        ]
    ]
]
\end{forest}\label{ex:spelloutwaarmee}

@@

\ex. \begin{forest} boom
[CP
    [\tsc{WP}
        [\tsc{WP}
            [\tsc{W}, roof]
        ]
        {\draw (.east) node[right]{⇒ \tit{w}}; }
        [\tsc{deixP}
            [\tsc{deixP}
                [\tsc{deix}, roof]
            ]
            {\draw (.east) node[right]{⇒ \tit{a}}; }
            [\tsc{accP}
               [\tsc{f}2]
               [\tsc{nom}P
                   [\tsc{f}1]
                   [ΣP
                       [Σ]
                       [\tsc{thingP}
                           [\tsc{thing}, roof]
                       ]
                   ]
               ]
            ]
            {\draw (.east) node[right]{⇒ \tit{'t}}; }
        ]
    ]
    [CP
        [...,roof]
    ]
]
\end{forest}

Only after the wh-element is moved to the left edge of the clause, case features higher than the accusative are merged.

When a dative feature is merged here, it will not be able to be spelled out. The only option to realize the features is to form to spawn a new derivation, to create a complex specifier. And that is exactly what forming the preposition \tit{met} `with' is.
